Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2025-06-27 Latex_files/Thesis/chapters/chapter 4/main.tex
updates 15/04/25
27/06/2025 chapters 1,3,4 and structural changes

\chapter{Complex Langevin dynamics of spherical dimers}
\label{chapter:langevin_dynamics}
Much of the calibration theory discussed in Chapter 2 
assumes that the target particle in question is a single 
sphere, one whose scattering and motion is easily computed. 
However, while working with dense colloidal suspensions, 
one often ends up trapping more than one sphere. Li and 
Arlt \cite{Li2008} studied the case of two microspheres 
trapped in a single OT and found that multiple trapped 
beads could be mistaken for a single trapped bead with 
altered trap stiffness (see \ref{sec:harmonic_traps}). 
Theoretical studies on the case of two trapped microspheres 
by Xu~\textit{et~al.} \cite{Xu2005} employed a ray-optics 
based model to show that the two trapped beads are brought 
into physical contact with each other by optical forces 
and they also calculated the axial equilibrium positions 
of the two trapped beads as a function of their size. 
Experiments in \cite{Praveen2016} confirmed that the two 
trapped beads indeed experience different trap stiffnesses 
in the vicinity of the same potential well.

There are further discussions looking into the dynamics of 
a whole host of asymmetrically shaped particles \cite{Loudet2014, 
ShengHua2005, Chetana2022}, their results all showing that 
predicting the behaviour an arbitrary shaped particle comes 
with great difficulty due to the fact that the optical force 
is dependent on a greater number of variables such as 
orientation and size factors.

With the initial goal of the project being to induce 
nucleation events via a rotating sphere, the aim of this 
chapter is to - in a limited capacity simulate and investigate 
the influence of a second particle being bound to our target 
sphere. The choice of a dimer, instead of an amorphous solid 
that might better represent a growing crystalline solid, 
allows us to consider how the dynamics of the aggregate change 
by varying a single parameter, namely the size ratio of the 
two spheres. Additionally, attempting to simulate an amorphous 
aggregate is rather difficult as calculating the optical force 
and torque is computationally slow. We build upon the works of 
Vigilante \textit{et al} \cite{Vigilante2020} to consider 
asymmetric spherical dimers and how varying size parameters 
alters the dynamics and additionally makes characterising their 
interactions within an optical trap more cumbersome.

In this chapter we will consider how the addition of a 
second sphere changes the trapping dynamics by introducing 
multiple equilibrium positions. These positions are 
dependent on both the size ratio of the dimer and its 
orientation relative to the trap. This is important to 
address, for while the the dimers' diffusive behaviour 
is the same, regardless of position and orientation, 
how it interacts with the trap may produce the wildly 
different results. Furthermore, we look at how our dimers 
interact with circularly polarised light, this is especially 
pertinent as the original plan was to utilise circularly 
polarised light to generate fluid flow in a supersaturated 
solution. The rotational motion of dimers in circularly 
polarised light goes against much of the established 
literature and suggests that the transfer of spin angular 
momentum is enabled by combining particles together. And 
lastly, we demonstrate how a quadrant photo diode, performs 
in characterising the actual interactions between a dimer 
and the optical trap. We do so to demonstrate that much 
of the dimer's trajectory information is either lost or 
poorly described by simple calibration techniques, making 
their motion difficult to characterise.

\subsection{Simulation Parameters}
As a paradigmatic example, consider a dimer suspended in 
water ($n_p = 1.59, n_m = 1.33$) located at the focus of 
a Gaussian beam (more specifically a Laguerre-Gaussian 
beam of mode $[0,0]$, see \eqref{eq:incident}), the beam 
is focused by a objective with numerical aperture of 1.2 
and is x polarised. The size ratio of the two spheres is 
given by $a_{I}/a_{II} = 2$ where $a_{I}$ is kept at $1\ 
\mu m$ unless specified otherwise; the dimer's orientation 
is given by a unit vector connecting the centres of both 
spheres (see \ref{sec:sim_parameters}). We define the 
'standard' orientation as being aligned with the direction 
of beam propagation direction - and therefore the 'inverted' 
orientation is defined when the dimer is orientated against 
the direction of beam orientation (see left hand side of 
figure~\ref{fig:paradigmatic}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Positional and Orientational dependence of Trapping forces}
\label{sec:eq_positions}
If we wanted to start from first principles and determine 
the trap strength on our target particle the first step 
would be to locate the equilibrium position relative to 
the trap focus. For a single sphere it is easy to enough 
to understand that its centre of mass will be drawn to 
focal point of the laser due to gradient forces, once there 
the force is analogous to a harmonic spring with a fixed 
trap stiffness (see fig.\ref{fig:harmonic_trap}). Now, if 
we consider instead a dimer, we now have two spheres both 
being drawn to the focus along by the same gradient force; 
in addition the scattering force is significantly more complex 
due to both spheres scattering the electromagnetic fields. 
This mutual scattering between individual spheres is what 
makes simulating spherical aggregates far more difficult 
compared to a single sphere, and even harder still to 
predict the position where the dimer's centre of mass is 
in equilibrium.

Because the scattering force is only significant in the 
direction of beam propagation \cite{Capitanio2002} we 
can assume there will only be a single potential well 
lying at the centre of the beam. Along the beam axis, 
such an assumption is no longer valid due to fact that 
spherical aggregates experience inter-particle scattering. 
This is a key reason for using \textit{mstm} as it 
accounts for that behaviour. The methodology for computing 
optical forces has been covered extensively for a number 
of different trapping conditions \cite{RanhaNeves2019}. 
So it is relatively easy to compute the trapping force 
and determine the equilibrium position by finding the 
position that minimises the net optical force and the 
local gradient is negative ($\delta F/\delta x < 0$) - 
we can assume that for a dielectric sphere the optical 
torque is negligible. For a dimer (or any arbitrary 
spherical aggregate), we now must consider both its 
position and orientation and find where the net optical 
force and torque are minimised.

After computing its $T$-matrix via \textit{mstm} and 
supplying that to \textit{ott} we compute the optical 
force exerted by a $50\ mW$ laser via \eqref{eq:optical_force} 
in the axial direction while the dimer is in its 'standard' 
orientation. As expected we see a single point where 
the dimer will be in equilibrium, the linear fit in 
fig.~\ref{fig:paradigmatic}(a) indicates a that the 
force can be modelled as a harmonic potential close to 
the equilibrium position ($F_z\approx-\kappa_z z$). 
The second point where the axial force goes to 0 cannot 
be considered as equilibrium position as the positive 
gradient indicates that the trap is unstable unless 
Brownian motion is ignored.

We repeated the same calculation but now while the 
dimer is in its 'inverted' orientation, instead of 
a single point where the optical force is minimised 
we see that there are instead two separate equilibrium 
positions, one above the focus and one below the 
focus. In this particular example the two positions 
are far enough apart that both can be considered as 
separate harmonic traps.    
\begin{figure}[h!]
	\centering
	\begin{subfigure}{.65\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=0.png}
		\caption{}
		\label{lam=2}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.25\linewidth}
		\centering
		\raisebox{60pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=0.png}}
		\label{large over small}
	\end{subfigure}
	\medskip
	\begin{subfigure}{.65\linewidth}
		\includegraphics[width=\linewidth]{lam=2_theta=180.png}
		\caption{}
		\label{lam=2_inverted}
	\end{subfigure}\hfill % <-- "\hfill"
	\begin{subfigure}{.25\linewidth}
		\centering
		\raisebox{40pt}[0pt][0pt]{\makebox{}\includegraphics[width=0.3\linewidth, keepaspectratio]{theta=180.png}}
		\label{small_over_large}
	\end{subfigure}
	\caption{Plots of force vs displacement of the 
		centre of mass of the dimer (µm) for the 
		case of a dimer of size ratio 2. (a) is the 
		case where the dimer is in its' 'standard' 
		orientation, where the dimer is trapped at 
		$z = 1.71\ \mu m$. (b) is the case where the 
		dimer is in its' 'inverted' orientation, the 
		dimer is trapped at two positions: $z = 1.20\ 
		\mu m$ \& $z=-1.63\ \mu m$. On the left are 
		renders to visualise the dimer orientation 
		are shown below each plot. The black lines 
		on each force-curve is a linear fit with the 
		slope being reported as the trap stiffness in 
		the legend.}
	\label{fig:paradigmatic}
\end{figure}

We can see that both equilibrium positions have comparable 
axial trap stiffness ($\kappa_z$), however the difference 
in the transverse trap stiffness ($\kappa_x$) is far more 
noticeable. The same dimer was trapped at each of the axial 
equilibrium positions and the transverse force was evaluated. 
While in all three cases the dimer can be trapped the linear
range where that would typically associated with a stable 
trap is far narrower in the 'standard' orientation compared 
to the 'inverted' cases. This highlights one of the challenges 
involved with studying asymmetric particles, even though its 
a simple enough process to trap them they maybe characterised 
very differently depending on their relative position and 
orientation towards the focus. This can have a significant 
impact on rheological studies - or attempting to probe any 
local property - as the variance in trap strength can result 
in large errors over repeated measurements. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{transverse_force.png}
	\caption{Plots of force vs displacement of the dimer's 
		centre of mass spheres, where a positive force 
		indicates the dimer is directed right on the x-axis, 
		and vice versa for a negative force. The same simulation 
		parameters are used here as in fig~\ref{fig:paradigmatic}
		(a) and (c). The blue curve representing the force 
		response for a dimer in its standard orientation, orange 
		being the inverted case, and green the same case but 
		placed below the focus.}
	\label{fig:transverse_force}
\end{figure}

\newpage
For completeness the harmonic traps were located for dimers
across a range of size ratios - from $a_{I}/a_{II} = 1$ to 
$a_{I}/a_{II}=10$ - while also recording the trap stiffness 
for each trap. The same simulation parameters are used here 
as for figures \ref{fig:paradigmatic} \& \ref{fig:transverse_force}. 
As shown in Fig.~\ref{fig:eq_positions} $a_{II}$ decrease 
the dimer begins to approximate a single homogenous sphere 
- at least in terms of location and trap strength. However, 
for intermediate sized dimers (between $a_{I}/a_{II} = 1.1$ 
to $a_{I}/a_{II}=4$), a second equilibrium position is found 
below the trapping focus. Previous work using the ray-optics 
model have confirmed even in the case that two spheres begin 
separated the electric field will align the particles as such 
that they make contact and are trapped together about a single 
trapping position \cite{Xu2005}. Furthermore it has been shown 
through proper manipulation of the Gaussian or Bessel beam 
modes that any number of trapping potentials can be developed 
\cite{Shahabadi2020} for nanoparticles. This result however, 
is the first example of an orientation dependent trapping 
situation using only a $TEM_00$ beam. Typical experimental 
arrangements cannot determine much information on the axial 
position of a trapped particle relative to the trap focus; 
this result indicates not only that dimers can be trapped 
in multiple axial positions but also their trapping behaviour 
is heavily dependent on said axial position. As such it is 
necessary that positional information in the z-axis can be 
elucidated if multiple spheres are trapped simultaneously. 
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers with varying 
		size ratio, dashed lines represent unstable traps whereas solid lines 
		are for stable equilibrium positions. (a) shows that dimers while in their
		'standard' orientation will always have a single equilibrium position. 
		(b) shows that when the same dimer is in its' 'inverted' orientation 
		can be trapped in two axial positions, one below the focus and one above 
		the focus.}
	\label{fig:eq_positions}
\end{figure}

\newpage
\subsection{Non-trivial equilibrium configurations}
\label{sec:off-axis}
Computing the equilibrium positions when a dimer is aligned 
with the electric field is relatively simple as the orientational 
torque is minimised (see Eq.\ref{eq:opt_torque}). Meaning once 
trapped the dimer is unlikely to change orientation enough to 
escape the trap. Regardless, that does not rule out the 
possibility that there is a stable configuration where the 
orientation not strictly vertical, in fact most experimental 
work with symmetric nano-dimers will trap them lying perpendicular 
to the beam direction \cite{Ahn2018, Reimann2018}. Unlike in Sec.~
\ref{sec:eq_positions} we cannot simply measure the optical force 
and torque as the parameter space is too large and determining if 
a particular position and orientation is stable is not clear based 
solely on force and torque measurements \cite{Bui2017}. Using the 
same simulation parameters as before we ran a number short 
simulations (total simulation time was $0.005$ s) with the laser 
power increased to $500\ mW$. Each simulation started with the 
dimer in a different starting position and orientation, due to the 
high laser power the dimers either escaped the trap or were stably 
trapped. The $z-\theta$ phase space - where $\theta$ is the angle 
between the direction of beam propagation and the dimer's orientation 
vector ($\theta=0^\circ$ is the 'standard' orientation) - can be 
divided into different regions depending on which equilibrium 
configuration is reached.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.67\linewidth}
		\includegraphics[width=\linewidth]{off_axis_trap.png}
	\end{subfigure}
	\begin{subfigure}{0.32\linewidth}
		\raisebox{0pt}[0pt][0pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
		\caption{}
	\end{subfigure}
	%\captionsetup{hangindent=2.35cm}
	\caption{Map of $z-\theta$ phase space using a dimer of
		size ratio 2 with a laser power of 500 mW ($\theta=0
		\circ$ is the 'standard' orientation and $\theta=180
		^\circ$ is the 'inverted' orientation). The stable 
		configurations are indicated by the larger circles 
		and the starting conditions are colour coded to match 
		the stable point they end up in. Right hand render 
		shows a dimer in its off-axis configuration.}
	\label{fig:off_axis}
\end{figure}

Interestingly while the trap strength of these off-axis traps 
are similar in magnitude to the vertically aligned dimers, 
but when the laser power is lowered (around $5\ mW$) the traps 
become metastable; after reaching its equilibrium configuration 
the particle behaves similarly to a typically trapped dimer but 
the trapping potential is small enough that the dimer escapes 
in as little time as less than a second or after nearly a full 
3 seconds. Running similar simulations but for vertical 
configurations sees the dimer remaining trapped, even after 30 
seconds of run time, indicating that the trapping potential is 
far greater than the thermal energy. This suggests that the 
reason this off-axis configuration is metastable is due to the 
increased rotational freedom. Similar configurations have been 
explored with ellipsoids; Zhu \textit{et al.} looked at the 
dynamics of various elliptical particles and found that regardless 
of shape or initial orientation the particle would tend towards
either a purely vertical or purely horizontal orientation 
\cite{Zhu2021}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Continuous rotational motion in circularly polarised light}
One aspect that has yet to be covered in depth with regards 
to spherical aggregates of any construction is their interaction 
with circularly polarised light. For homogenous spheres the 
optical torque is regarded as being negligible as the spin 
density cannot impart angular momentum while propagating in a 
homogenous medium. Dimers however, have been shown to experience 
an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018} 
while trapped in a circular polarised beam. In our simulations 
we found that dimers would rotate about their long axis when 
trapped in circularly polarised light. In this section we want 
to discuss how this behaviour is influenced by size, position, 
and orientation; and furthermore, we wish to address possible 
explanations for this behaviour, as none of the current
theories into optically induced rotation seem plausible.

\subsection{Polarisation Dependency on Dimer trajectory}
\label{sec:rot_pol}
In their paper Vigilante \textit{et al} attribute the rotational motion 
to the anisotropic shape of the dimer \cite{Vigilante2020}. Anisotropic 
scattering is a viable theory for describing optical rotation, however
there is usually an optical axis about which rotation occurs. For a dimer
this is typically the orientation vector that passes through the centres 
of both spheres \cite{Ahn2018, Reimann2018, Bruce2020}. There have been 
cases where the particle's cross sectional shape is engineered to scatter
light in one particular direction \cite{Higurashi1994}, but if this was
the case then the dimer should rotate when illuminated by any polarisation
of light.

>>>>

\chapter{Detection of rotational spherical aggregate rotational dynamics}
\textit{\textbf{Add overview here}} \\
\label{chapter:simulated_detection}

\section{Overview}
In this chapter we go about defining an optical fibre detection 
system from a mathematical perspective. Before demonstrating how 
it can be used to instantaneously predict orientational information 
on a trapped particle. We then consider common factors of error in the 
characterisation process, such as signal error, and incorrect particle 
sizing. Both of which have a significant impact on the performance 
of the model, but by utilising Bayesian inference, and time averaging 
we can better refine our prediction to a satisfactory degree.

\section{Introduction}
As outlined in the end of Chapter 3, one of the difficulties 
in characterising interactions with asymmetric objects is 
the coupled motion between translation and rotation. In order
to characterise the optical trap a position detection system 
such as a lateral effect photodiode or a quadrant photo diode
(see fig.\ref{fig:position_detectors}). Typically, a position 
detection system assumes that there exists a linear relationship 
between the a particle's displacement and the detected signal.
However, this is not always the case as demonstrated in chapter
4; as dimer's demonstrate non-harmonic trapping beyond a certain 
displacement. In addition, the rotational effects of the trapped
dimer become a more significant factor when characterising the 
trap strength. This can have unintended effects when it comes to
areas of research that require a precise understanding of the 
hydrodynamic behaviour of particles. As such there has been a 
concerted effort to either eliminate rotational motion, or describe
how the rotational behaviour influences the trajectory reported 
by the position detector.

In this chapter we build upon the work of \cite{BarZiv1997} to 
determine the instantaneous orientation of the aggregate. Using 
static light scattering we can map the outputted signal from a 
optically trapped dimer to its expected orientation in real time. 
Thus allowing us to estimate the optical torque being applied to 
the dimer by the trapping beam. As demonstrated in chapter 3, 
the influence of a circularly polarised beam can result in 
gyroscopic precession for specifically sized dimers.

\section{Monitoring Stochastic rotational motion using static 
		light scattering}
Part of the optical fibre arrangement borrows code from the 
optical tweezer simulations discussed in Chapter 2 and 4. 
Rather than repeat the simulation details we will instead 
only focus on the specifics of the optical fibre scattering.

\subsection{Coordinate System}
Consider a particle (either a sphere or dimer) trapped by
a Gaussian beam, with its focal point being set at [0,0,0]. 
At the same time an optical fibre directs a plane wave that 
is incident on the trapped particle scattering light in all 
directions. The probe beam is assumed to be x-polarised and 
propagating in the +z direction, this is a $90^{\circ}$ 
rotation from the previous simulations (see fig\emph{X.X}).

\subsection{Dimer}
The dimer is defined by two spheres with refractive indices
of 1.59 and suspended in water ($n_{med} = 1.33$). \textit{MSTM}
requires a target origin for setting the scattering expansion, 
there is no agreed upon solution for where the target origin 
should be relative to the sphere positions. In chapter 4 we showed
that altering the target origin would drastically alter the 
tmatrix produced. We choose (somewhat arbitrary) to use 
the dimer's centre of diffusion which is located $\zeta a_I$ from 
the centre of the larger sphere. For a discussion see \emph{X.X}.

\subsection{Scattering and Detection}
Each detector is placed some distance $\bf r$ from the origin, 
detector positions are defined by angles $\theta$ \& $\phi$. With 
$\theta=\phi=0$ being directly in front of the probe beam. The 
position of each detector is given by:
\begin{align}
	\begin{pmatrix}
		x_{fiber} \\ y_{fiber} \\ z_{fiber}
	\end{pmatrix} = 
	\begin{pmatrix}
		rcos(\phi)sin(\theta) \\ rsin(\phi)sin(\theta) \\ rcos(\theta)]
	\end{pmatrix}
\end{align}
Where r is the distance between the detector and the target origin. 
In an ideal situation the surface of the detector is oriented towards 
the target origin perfectly. Pixels can be thought of as points lying 
on the surface of the detector, each point has its own position in real
space. Using \textit{mstm} we can find the components of the scattering 
matrix at each point and compute the intensity of the electric field via:

\begin{align}
	\begin{pmatrix}
		I_s \\ Q_s \\ U_s \\ V_s
	\end{pmatrix} = 
	\begin{pmatrix}
		S_{11} & S_{12} & S_{13} & S_{14} \\
		S_{21} & S_{22} & S_{23} & S_{24} \\
		S_{31} & S_{32} & S_{33} & S_{34} \\
		S_{41} & S_{42} & S_{43} & S_{44} \\
	\end{pmatrix}
	\begin{pmatrix}
		I_i \\ Q_i \\ U_i \\ V_i
	\end{pmatrix}
\end{align}
Where:
\begin{align}
	I_s &= E_{x,\ scat}^2+E_{y,\ scat}^2  \label{eq:intensity} 
	\\
	\rightarrow I_s &= S_{11}I_i+S_{12}Q_i+S_{13}U_i+S_{14}V_i
\end{align}

By computing the scattering at each point on the detectors surface 
we can compute the average scattering over the detector. From a
experimental perspective the distance from the particle to the 
detectors must be considered. The near field scattering can be 
computed with relative ease using \textit{mstm} as shown in 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{near_field_intensity_side_profile.png}
	\caption{Intensity distribution (as defined by \eqref{eq:intensity}) 
		from a symmetric dimer when irradiated by the probe beam 
		propagating in the $k\hat{z}$ direction (viewed in the 
		z-y plane). Axis are given in dimensionless units of 
		$k\hat{z}$ and $k\hat{y}$.}
	\label{fig:nf_scattering}
\end{figure}

While \textit{mstm} does provide the amplitude matrix it is not
scaled to account for distance from the target to the detector.
We amend this by simply scaling according to the inverse square 
law. As the detector is moved further from the target the 
intensity distribution across the entire detector will decrease
substantially. Because of the fall off it is more computationally
efficient to instead consider only the central pixel of the 
detector rather than across the entire surface. If we consider 
the difference between the average intensity and the intensity 
of the central point we can see that after not even $20 \mu$ 
away from the target particle the average intensity across the 
surface is within $0.1\%$ of the central pixel. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{fall_off_with_distance.png}
	\caption{Normalised difference between the central pixel intensity
	$I(\theta, \phi)$ and the average pixel intensity over the surface
	of the optical fibre $\langle I \rangle$ as a function of distance.
	4 different optical detectors were tested (spherical coordinates 
	are given in the legend). While the choice of coordinates is 
	arbitrary they all demonstrate the same exponential decay. Dashed
	lines indicate the difference after $20\mu m$ from the trap focus,
	in most practical cases getting an optical fibre any closer would
	disrupt the trap stability.}
\end{figure}

Given the trapping beam's beam waist is approximately $0.54 \mu m$,
a optical fibre would need to be a minimum of $150 \mu m$ from the 
trap centre in order to prevent noise from the trapping laser itself.
At such a distance the variation in the signal is so small that it is 
easier to evaluate the intensity at the centre of the detector instead
of across the entire surface. This is how the instantaneous orientation
is predicted later on in the chapter.

\section{Interpretation of scattering data into orientation estimates}
We now discuss the methodology for computing the instantaneous 
orientation of an optically trapped particle based on scattering
detected by a finite grouping of optical fibres.

Consider a dimer in the optical trap (Fig.~\ref{fig:dimer}a), 
we can define at any point in time a unit vector $\hat{s}$ 
pointing from the centre of the larger sphere to the centre 
of the smaller sphere. A plane wave probe beam incident, 
is incident on the dimer, generating a scattering pattern
dependent on the dimer's orientation $I(\hat{\bf s}, 
\theta, \phi)$ which is computed using \textit{mstm} 
\cite{I.Mishchenko1996}. To represent the experimental set 
up consisting of a set of optical fibres recording scattered 
light, we choose four sets of spherical angles [$(\theta_1,\ 
\phi_1), \ (\theta_2,\ \phi_2), \ (\theta_3, \ \phi_3), \ 
(\theta_4,\ \phi_4)$] and record the calculated intensity at 
each angle $I(\hat{s},\ \theta_k,\ \phi_k)$. 
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig2a.png}
	\end{subfigure}
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig2b.png}
	\end{subfigure}
	\caption{(a)Example dimer in orientation $\hat{\bf s}$, (b) 30 Reference orientations represented by vectors pointing from [0,0,0] to each point}
	\label{fig:dimer}
\end{figure}

Our goal is to determine the orientation of the trapped dimer 
based on the measured intensity $I(\hat{n}, \ \theta_k,\ \phi_k)$. 
Rather than aim an exact estimate of the dimer's orientation, 
for the purposes of interpretation of the scattering and 
optimisation of the measurement setup it is more convenient 
to discretize the possible orientation space into a number 
of possible reference orientations. We can then use each as 
'classification categories' in a neural network methodology 
to map scattering data to orientation (see below for further 
discussion). Here we choose $\textit{n}_{ref} \ = \ 30$ 
reference orientations $\hat{\bf n}_{\alpha}$  evenly 
distributed on a unit sphere \cite{Reyuthor2006} 
(Figure~\ref{fig:dimer}b) leading to a maximum nearest 
neighbour spacing between two neighbouring reference 
orientations of 0.895 radians.

Using \textit{mstm} we compute the raw intensities (see 
\eqref{eq:intensity}) at each of the measurement angles 
that would be generated by a dimer in each reference 
orientation, $I_s(\hat{\bf n}_{\alpha}, \theta_k)$. While 
the number and position of detection fibres is technically 
arbitrary there are several constraining factors that limit 
our ability to infer useful information from the trapped 
object, see Section~\ref{sec:brownian} for a detailed 
breakdown of our choice of detection angles. In order for 
a neural network to properly weight each signal the raw 
intensities are normalized according to:
\begin{align}
	\label{eq:scale}
	y_k(\hat{\bf n}_\alpha)
	= 
	\frac{I(\hat{\bf n}_\alpha, \theta_k) - \langle I(\hat{\bf n},\theta_k) \rangle } 
	{\langle I^2(\hat{\bf n},\theta_k) \rangle -\langle I(\hat{\bf n}, \theta_k)\rangle^2}
\end{align}

where the denominator is simply the standard deviation across 
the set of values $I(\hat{\bf n}, \theta_k)$. Note that the 
collected scattering signals are not necessarily simply 
related to their associated reference orientations: as is well 
known from such examples of the inverse scattering problem. 
While it is trivial to compute the light scattering pattern 
for any given particle with any particular characteristic (i.e. 
size, shape, or orientation), inferring the light scattering 
from a unknown particle to determine said characteristic is 
incredibly difficult due to complex mapping between scattering 
and said characteristic. Even if the orientation space is 
divided evenly between reference orientations the subsequent 
signal space ends up being appearing mixed making simple 
comparisons of signals useless for inferring information on the 
particle.

Fig~\ref{fig:mixing} shows two clusters of orientation 
vectors and their respective signals collected from 3 detectors
- the points have been coloured based on their proximity 
to the centre of their respective cluster. While the 
orientation space appears tightly packed and ordered the 
signal space quickly spreads out in an asymmetric fashion. 
Furthermore as seen in Fig~\ref{fig:mixing}b the signal 
mapping can intersect itself which only further increases 
the complexity. While in some instances the mapping between 
one reference orientation and another is discrete, in other 
instances the mapping becomes far more complex to discern.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.74\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig3a.png}
	\end{subfigure}
	\begin{subfigure}{0.74\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{fig3b.png}
	\end{subfigure}
	\caption{(a) Distribution of orientation vectors and (b) their respective scattering signals. Points are coloured according to their distance from the centre of each cluster (red points centred around [$0.00, 0.00, 1.00$], blue points centred at [$0.71, 0.00, 0.71$])}
	\label{fig:mixing}
\end{figure}

Nevertheless, at least where the uncertainty in signal measurements 
is low, we can predict the orientation from the scattering scattering 
by utilising computational techniques such as neural networks. We 
thus utilised the Python machine learning program \textit{scikit-learn} \cite{Pedregosa2011} to build a neural network for identifying the 
dimer's orientation from its light scattering signal. The network 
was trained by generating a database of random orientation vectors, 
calculating the corresponding light scattering signals, and then 
using the network to estimate the probability of a given signal coming 
from a dimer in a given reference orientation. The network's loss 
function was evaluated and used to improve the estimation, the 
network being trained until the improvement in the loss function 
was less than 0.0001.

Importantly, the estimation provided by the neural network can be 
improved further by accounting for any prior information we know 
about the dimer, utilising Bayesian inference to update the neural 
network's estimation: 
\begin{align}
	\label{eq:bayes}
	p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}))
	&=
	\frac{p(y_k(\hat{\bf s})|\hat{\bf n}_\alpha)
		p(\hat{\bf n}_\alpha)}{p(y_k(\hat{\bf s}))}
\end{align}

where $p(\hat{\bf n}_\alpha)$ and $p(y_k(\hat{\bf s}))$ are the prior 
estimates of the distributions of particle orientations and 
instantaneous signals, respectively. \textit{Without} any prior 
evidence we must assume that the orientation prior of the dimer 
$p(\hat{\bf n}_{\alpha})$ is uniform. However, inference about the 
dimer's possible current orientation from knowledge of previous 
measurements can be used to inform our estimate of $p(\hat{\bf 
n}_{\alpha})$. The latter prior $p(y_k(\hat{s}))$ is the probability of 
measuring a signal ($y_1$, $y_2$, $y_3$). This is given by taking 
the discrete integral over the collection of reference orientations:
\begin{align}
	p(y_k(\hat{\bf s}))
	=
	\sum_{\alpha=1}^{n_{\rm ref}}
	p(y_1, y_2, y_3, y_4|\hat{\bf n}_\alpha)
	p(\hat{\bf n}_\alpha)
\end{align}

From \eqref{eq:bayes} we obtain the key result, a mass probability 
distribution denoting the probability that our dimer is in 
orientation $\hat{\bf n}_{\alpha}$ given a measured signal ($y_1$, 
$y_2$, $y_3$), \textit{i.e.} an estimated mapping from scattering 
measurement to orientation estimate.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Calculation of error}
\label{sec:divergence}
Using \eqref{eq:bayes} we get a snapshot estimation of the instantaneous
orientation of the trapped dimer based on a single measurement from a 
collection of detector signals. Of course in reality these detectors 
will be returning a discrete time series representative of the dimer's
trajectory within the optical trap.

Using the simulation code from \cite{Vigilante2020} we can replicate 
this trajectory and furthermore use \textit{mstm}to calculate the far field 
intensity produced by said dimer. At each time step we measure the signal
produced at each measurement angle and use our neural network to calculate
$p(y_k(\hat{\bf s})|\hat{\bf n}_\alpha, t)$.

Evaluating how well Eq.~\eqref{eq:bayes} performs in predicting $\hat{\bf n}$
is difficult because we have discretised the orientation space meaning 
there will always be some degree of inaccuracy in our prediction. The Kullback-Liebler divergence method provides a measure of the relative
entropy change between two probability distributions:
\begin{align}
	D_{KL}(P(t)\parallel Q(t)) = \sum P(t) \log_2\left[\frac{P(t)}{Q(t)}\right] 
	\label{eq:D_KL}
\end{align}

Where $P(t)$ and $Q(t)$ are two probability distributions we are
comparing, often it is common to use a base 2 log as this converts 
the entropy to base 2. By relative entropy we are talking about 
the useful information gained/lost by switching from $P(t)$ to 
$Q(t)$. In our case $Q(t)$ is simply the distribution given by
Eq.~\eqref{eq:bayes} and $P(t)$ is an idealised distribution, 
one where the model is accurate with 100\% certainty.
\begin{align}
	\label{eq:best}
	p_{best} = P(t) =
	\begin{cases}
		1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
		0 & \text{when $\hat{\bf n}_\alpha$ $\neq$ $\hat{\bf n}_{best}$}
	\end{cases}
\end{align}

In reality the distribution from Eq.~\eqref{eq:bayes} will assign 
some non-zero probability to every reference orientation. Therefore 
the relative entropy for a single measurement is given as:
\begin{align}
	D_{KL}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t))
	= \log_2 \left[\frac{1}{p(\hat{\bf n}_\alpha| y_k(\hat{\bf s}), t)}
	\right]
	\label{eq:kullback}
\end{align}

The relative entropy is a measure if how spread out our distribution
is. With larger values of $D_{KL}$ indicating that our prediction is 
spread out so the neural net is unsure about $\hat{\bf n}_{best}$. 
Whereas a low value of $D_{KL}$ means that our model is more confident 
in $\hat{\bf n}_{best}$. The only draw back being that the relative 
entropy does nothing to indicate accuracy, only how closely our 
distribution matches our ideal result. This will be better 
exemplified later in \ref{sec:test}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Brownian Simulation}
\label{sec:brownian}
